Newly Left-Behind Children's Educational Performance Drops Slightly After Parental Migration

### 2022-07-05 MScThesisText.tex
Final

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Newly Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated},
  pdfauthor={Jiancheng Gu1},
  pdflang={en-EN},
  pdfkeywords={labor migration, left-behind children, educational performance},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{Newly Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated}
\author{Jiancheng Gu\textsuperscript{1}}
\date{}

@@

\abstract{%
Millions of children worldwide are ``left behind'' at home communities as their parents migrate away for work.
Parents' labor migration increases household income but decreases parental care, thereby exerting mixed influences on children's educational attainment.
To what extent do different arrangements of parental migration affect children's educational performance in the short term?
Studying the case of China, I obtained a sample of four thousand junior high school students from a nationally representative two-wave panel survey.
I differentiated household migration arrangements of any one parent, of mother only, of father only, and of both parents.
I measured children's schooling performance by cognitive test scores and academic exam grades.
For causal inference, I built difference-in-differences models with propensity score matching.
Results show that the newly left-behind children's performance at school dropped slightly compared to that of non-left-behind peers.
Mother-only migration is negatively associated with children's academic grades and cognitive abilities.
Father-only migration is negatively associated with children's cognitive abilities, but its magnitude is greater than mother-only migration.
}

@@

A challenge arises in estimating the effect of parent's migration on children's educational achievement. Experiments can unveil causal relations. It is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. It requires that the cause br on par with random assignment to participants, conditional on identification assumptions (Cunningham, 2021).

In this thesis, my research question is: To what extent do different arrangements of parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyze data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and applied propensity score matching for robustness checks. The outcome was measured with children's cognitive test score and academic exam grade. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

@@

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers. They form the backbone of essential industries, such as retail, delivery, transportation, and manufacturing. The 2015 one-percent National Population Sample Survey estimated that 68.7 million Chinese children were left behind (NBS, 2022; NBS et al., 2017).

The internal migration stemmed from the neoliberal economic regime. Scholars like Duckett (2020) and Zhan (2020) argued that China conforms to the neoliberal trend, extending the logic of the market to other domains of human life. In the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. Simultaneously, many people have left the countryside for the city. The regional inequality needs social policy to amend. But China's welfare program, as Qian (2021) argued, is fragmented and stratified. Those who reap the greatest benefits are public servants, urban dwellers, and formally employed workers. As rural migrants arrive in the city, they find themselves with limited access to social welfare, such as health care and children's education (Qian, 2021). The welfare system's exclusion of rural migrants means that the state shifts its responsibility to individuals, resembling the neoliberal governmentality (Zhan, 2020).

The institutionalized discrimination against rural migrants finds its roots in China's ``multilevel citizenship'' system. As Vortherms (2021) explained, the constituents of citizenship -- ``membership rules and rights entitlements'' -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through the \emph{hukou}, or household registration system:

@@

How does China counter the issue of labor migration and left-behind children? According to X. Wang (2020) and Zhang (2018), the government has gradually liberalized the multilevel citizenship scheme. It has introduced residence permit for internal migrants as the basis for social entitlements. First-tier cities like Beijing and Shanghai have brought forth a point-based system to grant hukou to newcomers. Criteria include in-demand talents and skills, educational and professional qualifications, and consecutive years of employment and tax-payment. This parallels how European states differentially incorporate foreign immigrants into national welfare regimes (Zhang, 2018). Another policy response, as stated by Roberts and Hannum (2018), is the centralized boarding schools. The Ministry of Education promotes boarding schools in the rural county-seats. These schools serve as a cost-efficient way to enroll students in surrounding villages and to look after the left-behind children. In reality, such schools ``are often ill-equipped to ensure the healthy upbringing and psychological wellbeing of children'' (p.~6).

Social scientists have researched extensively about China's left-behind children, but the current literature still has three major shortcomings. The first is that most literature focuses on children left behind in rural areas while neglecting the urban counterparts. The National Bureau of Statistics stated that the number of urban left-behind children had been growing and expected that the trend would continue (NBS et al., 2017). My sample includes urban left-behind children as well. The second shortcoming lies in sampling and research design. Some studies suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one to a few towns (Y. Liu et al., 2021; Shu, 2021; F. Yang et al., 2022). Others only estimated correlation using cross-sectional data (Akezhuoli et al., 2022; Huang et al., 2020; Jin et al., 2020). These studies could have employed causal inference techniques like fixed effect, matching, or inverse probability of treatment weighting. The third shortcoming lies in the measurement of outcome. Chang et al. (2019) and L. Wang et al. (2019) measured students' educational performance with the standardized test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. I included the test results of multiple subjects to measure comprehensive academic achievement.

@@

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, ``immediately and directly convertible into money and may be institutionalized in the form of property rights'' (2002, p. 16). Economic capital is transferable to human capital, which is the stock of ``skills, talents, health, and expertise'' in people that enhance their productivity (Botev et al., 2019; Goldin, 2016). D. Yang (2008) noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational expenditure. Cebotari (2020) developed a fixed effect model with three-wave panel data collected in Ghana. He found that stay-behind children's education benefits from ``the steady inflow of both in-kind and monetary remittances'' contributed by migrant parents. In China, Chang et al. (2019) interviewed migrant workers and their family members. Oftentimes, they found that labor migration presented the only way to sustain their children's education and livelihood. Peng (2018) interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, ``good mothering'' primarily means offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the ``stable environments that promote health and adequate nutrition, protect from threats, and provide opportunities for learning and responsive, emotionally supportive and developmentally enriching relationships'' (Black et al., 2021, p. 1). Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents (Wen \& Xie, 2019). Hong talked to a group of left-behind children at a rural school (Hong \& Fuller, 2019). These students expressed the utter sense of loneliness ``with no advocate, no support and with no one looking out for them'' (p.~13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p.~14). Similarly, Murphy (2022) talked to left-behind adolescents at a village. Some of them ``embraced the `left-behind' label'' because it captured their need for more parental care (p.~185), ``especially companionship and affection in daily life'' (p.~186).

Given the two opposing effects, researchers have obtained mixed results on how parental migration impacts children's education. Botezat and Pfeiffer (2020) adopted past immigration rate as an instrumental variable. They found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen in Romania. L. Wang et al. (2019) collected data from junior high school students in central China. The researchers' two-way fixed-effect model showed no effect of parental on left-behind children's achievement in mathematical test. Studying four Chinese provinces, Brauw and Giles (2017) showed that as ample migration opportunities appeared in villages, enrollment to senior high school decreased. Nguyen (2016) conducted fixed-effect regression and demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight.

Parental migration immediately disrupts family structure, which in turn lowers academic performance of children. It takes longer time for parental migration to generate resources that improve children's educational attainment. I propose the first hypothesis:

@@

Researchers have studied whether the impact differ between mother's migration and father's migration. Xu et al. (2019) adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had ``little or no association with negative outcomes'' on children's academic and cognitive performances. Only-mother absence, however, showed a ``strong association with negative outcomes'' (p.~1646). Chen et al. (2019) conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was ``negatively associated with children's social competence and academic performance,'' while father migration had no direct effects on the two outcomes (pp.~860-861).

@@

I obtained the data from the China Education Panel Survey (CEPS) (NSRC, n.d.). It is a nationally representative, school-based survey featuring junior high school students. Currently, two waves of data are available. The CEPS team operates under the National Survey Research Center at the Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The baseline survey (wave one) took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. These students were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties (hereafter `counties'). The follow-up survey (wave two) in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed 9449 of the seventh grade cohort. I further restricted the sample to students living with both parents at the original domicile at the baseline survey. To tackle missing values, I adopted list-wise deletion for the dependent, independent, and control variables. The final sample size amounts to 4062.

@@

The key independent variable, the parental migration status, was identified by two items in the parent survey. In both the baseline and the follow-up surveys, parents specified the family members living in the same household at the time. I constructed four treatment dummy variables measuring migration arrangements: any parents migrated, only father migrated, only mother migrated, and both parents migrated. These arrangements are mutually exclusive. The control group consisted of parents staying with their children throughout at both the baseline and the follow-up surveys. I then removed the observations that reported parental divorce or death. These type of parental absence are irrelevant to migration.

The control variables on children's level include age, gender, ethnicity, self-rated health, hukou type, number of siblings, whether the child had skipped grade or repeated grade, and the level of education that the child aspires to achieve. At the household level, I controlled the levels of education of father and mother, the parental expectation on the child's level of education, whether the parent has a white collar job, whether the household has internet access, the perceived household financial conditions, and the number of extracurricular books at home. The lists of covariates are in Tables 3-5 in the Results section.

@@

I built a two-way fixed effect model in a difference-in-differences (DID) design, following Bai and colleagues' (2018; 2020) approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. I first specified an unrestricted and unadjusted model: \[\Delta score_{i,j} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i;1} + \lambda \cdot S_{i} + \varepsilon_{i,j}\] For student \(i\) in school \(j\), \(\Delta score_{i,j}\) is the change in test score between baseline and follow-up surveys, \(migr_{i}\) is the treatment dummy variable, \(score_{i,j;1}\) is the test score at baseline, and \(S_{j}\) is the school fixed-effect. School fixed-effect can account for differences due to time-invariant factors at the school and county levels, as students nested in the same school are also in the same county. An alternative is to use a class fixed-effect to control for time-invariant factors in the classroom. However, about one-sixth of the students were reassigned into new classes at the end of baseline academic year. As the reassignment of classroom will complicate the model, I ruled out the classroom fixed-effect.

The DID model is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed effect model (Huntington-Klein, 2022). The standard errors were clustered at the classroom level.

In addition, I present an unrestricted and adjusted DID model: \[\Delta score_{i,j} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i;1} + \lambda \cdot S_{i} + \theta \cdot X_{i} + \varepsilon_{i,j}\] where \(X_{i}\) is the vector of covariates capturing the characteristics of children, their parents, and their households. The control variables were measured in the baseline survey. This version of model lifts the restriction that covariates from the baseline survey would be associated with a coefficient that equals one.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Newly Left-Behind Children's Educational Performance Drops Slightly After Parental Migration},
  pdfauthor={Jiancheng Gu1},
  pdflang={en-EN},
  pdfkeywords={labor migration, left-behind children, educational performance},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{Newly Left-Behind Children's Educational Performance Drops Slightly After Parental Migration}
\author{Jiancheng Gu\textsuperscript{1}}
\date{}

@@

\abstract{%
Millions of children worldwide are ``left behind'' at home communities as their parents migrate away for work.
Parents' labor migration increases household income but decreases parental care, thereby exerting mixed influences on child development outcomes like educational attainment.
To what extent do different arrangements of parental migration affect children's educational performance in the short term?
Studying the case of China, I obtained a sample of four thousand junior high school students from a nationally representative two-wave panel survey.
I differentiated household migration arrangements of any one parent, of mother only, of father only, and of both parents.
I measured children's schooling performance by cognitive test scores and academic exam grades.
For causal inference, I built difference-in-differences models with propensity score matching.
Results show that the newly left-behind children's performance at school dropped slightly compared to that of non-left-behind peers.
Mother-only migration is negatively associated with children's academic grades and cognitive abilities.
Father-only migration is negatively associated with children's cognitive abilities, but its magnitude is greater than mother-only migration.
}

@@

A challenge arises in estimating the effect of parent's migration on children's educational achievement. Experiments can unveil causal relations. It is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. It requires that the cause be on par with random assignment to participants, conditional on identification assumptions (Cunningham, 2021).

My research question is: To what extent do different arrangements of parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyze data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and apply propensity score matching for robustness checks. The outcome is measured with children's cognitive test score and academic exam grade. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

@@

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers. They form the backbone of essential industries, such as retail, transportation, and manufacturing. The 2015 one-percent National Population Sample Survey estimated that 68.7 million Chinese children were left behind (NBS, 2022; NBS et al., 2017).

The internal migration stemmed from the neoliberal economic regime. Scholars like Duckett (2020) and Zhan (2020) argued that China conforms to the neoliberal trend, extending the logic of the market to other domains of human life. In the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. Simultaneously, many people have left the countryside for the city. The regional inequality needs social policy to amend. But China's welfare program, as Qian (2021) argued, is fragmented and stratified. People who reap the greatest benefits are public servants, urban dwellers, and formally employed workers. As rural migrants arrive in the city, they find themselves with limited access to social welfare, such as health care and children's education (Qian, 2021). The welfare system's exclusion of rural migrants means that the state shifts its responsibility to individuals, resembling the neoliberal governmentality (Zhan, 2020).

The institutional discrimination against labor migrants finds its roots in China's ``multilevel citizenship'' system. As Vortherms (2021) explained, the constituents of citizenship -- ``membership rules and rights entitlements'' -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through \emph{hukou}, or the household registration system:

@@

How does China counter the issue of labor migration and left-behind children? According to X. Wang (2020) and Zhang (2018), the government has gradually liberalized the multilevel citizenship scheme. It has introduced residence permit for internal migrants as the basis for social entitlements. First-tier cities like Beijing and Shanghai have brought forth a point-based system to grant hukou to newcomers. The system's evaluation criteria include in-demand talents and skills, educational and professional qualifications, and consecutive years of employment and tax-payment. This parallels how European states differentially incorporate foreign immigrants into national welfare regimes (Zhang, 2018). Another policy response, as stated by Roberts and Hannum (2018), is the centralized boarding school. The Ministry of Education promotes boarding schools in the rural county-seats. These schools serve as a cost-effective way to enroll students in surrounding villages and to look after the left-behind children. In reality, such schools ``are often ill-equipped to ensure the healthy upbringing and psychological wellbeing of children'' (p.~6).

Social scientists have researched extensively about China's left-behind children, but the current literature still has three major shortcomings. The first is that most literature focuses on children left behind in rural areas while neglecting the urban counterparts. The National Bureau of Statistics stated that the number of urban left-behind children had been growing and expected that the trend would continue (NBS et al., 2017). The second shortcoming lies in sampling and research design. Some studies suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one to a few towns (Y. Liu et al., 2021; Shu, 2021; F. Yang et al., 2022). Others only estimated correlation using cross-sectional data (Akezhuoli et al., 2022; Huang et al., 2020; Jin et al., 2020). These studies could have employed causal inference techniques like fixed effect, matching, or inverse probability of treatment weighting. The third shortcoming lies in the measurement of outcome. Chang et al. (2019) and L. Wang et al. (2019) measured students' educational performance with the standardized test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. My sample includes urban left-behind children as well. I included the test results of multiple subjects to measure comprehensive academic achievement.

@@

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, ``immediately and directly convertible into money and may be institutionalized in the form of property rights'' (2002, p. 16). Economic capital is transferable to human capital, which is the stock of ``skills, talents, health, and expertise'' in people that enhance their productivity (Botev et al., 2019; Goldin, 2016). D. Yang (2008) noted that in the Philippines, a sharp depreciation of peso increased the amount of remittances from migrants working abroad. This subsequently boosted children's school enrollment and educational expenditure. Cebotari (2020) developed a fixed effect model with longitudinal data collected in Ghana. He found that stay-behind children's education benefited from ``the steady inflow of both in-kind and monetary remittances'' contributed by migrant parents. In China, Chang et al. (2019) interviewed migrant workers and their family members. Oftentimes, labor migration presented the only way to sustain their children's education and livelihood. Peng (2018) interviewed first-generation migrant mothers whose initial migration dated back to the 1990s. To them, ``good mothering'' primarily meant offering economic support for their children in education, which can help their children climb the ladder of success.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the ``stable environments that promote health and adequate nutrition, protect from threats \ldots{} opportunities for learning,'' and ``responsive, emotionally supportive and developmentally enriching relationships'' (Black et al., 2021, p. 1). Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents (Wen \& Xie, 2019). Hong talked to a group of left-behind children at a rural school (Hong \& Fuller, 2019). These students expressed the utter sense of loneliness ``with no advocate, no support and with no one looking out for them'' (p.~13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p.~14). Similarly, Murphy (2022) talked to left-behind adolescents at a village. Some of them ``embraced the `left-behind' label'' because it captured their need for more parental care (p.~185), ``especially companionship and affection in daily life'' (p.~186).

Given the two opposing effects, researchers have obtained mixed results on how parental migration impacts children's education. Botezat and Pfeiffer (2020) adopted past immigration rate as an instrumental variable. They found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen in Romania. The team of L. Wang et al. (2019) collected data from junior high school students in central China. Their two-way fixed-effect model showed no significant effect of parental migration on left-behind children's achievement in mathematical test. Nguyen (2016) conducted fixed-effect regression and demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight.

Parental migration immediately disrupts family structure, which in turn lowers academic performance of children. It takes longer time for labor migration to generate resources that improve children's educational attainment. I propose the first hypothesis:

@@

Researchers have studied whether the impact differ between mother's migration and father's migration. Xu et al. (2019) adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had ``little or no association with negative outcomes'' on children's academic and cognitive performances. Only-mother absence, however, showed a ``strong association with negative outcomes'' (p.~1646). Chen et al. (2019) conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that maternal migration was ``negatively associated with children's social competence and academic performance,'' while paternal migration had no direct effects on the two outcomes (pp.~860-861).

@@

I obtained the data from the China Education Panel Survey (CEPS) (NSRC, n.d.). It is a nationally representative, school-based survey featuring junior high school students. Currently, two waves of data are available. The CEPS team operates under the National Survey Research Center at the Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The wave one baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. They were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties (hereafter `counties'). The wave two follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed 10279 of the seventh grade cohort. I further restricted the sample to students living with both parents at the original domicile (place of hukou registry) at the baseline survey. To tackle missing values, I adopted list-wise deletion for the dependent, independent, and control variables. The final sample size amounts to 4062.

@@

The key independent variable, the parental migration status, was identified by two items in the parent survey. In both the baseline and the follow-up surveys, parents specified the family members living in the same household at the time. I constructed four treatment dummy variables measuring migration arrangements: any parents migrated, only father migrated, only mother migrated, and both parents migrated. The latter three arrangements are mutually exclusive. The control group consisted of parents staying with their children throughout the two waves of survey. I then removed the observations that reported parental divorce or death. These type of parental absence are irrelevant to migration.

The control variables at children's level included age, gender, self-rated health, hukou type, number of siblings, whether the child is ethnic minority of Han Chinese, whether the child had skipped grade or repeated grade, and the level of education that the child aspires to achieve. At the household level, I controlled the levels of education of father and mother, the parental expectation on the child's level of education, whether the parent has a white collar job, whether the household has internet access, the perceived household financial conditions, and the number of extracurricular books at home. The lists of covariates are in Table~\ref{tab:tab-ctvar-dit} -- Table~\ref{tab:tab-ctvar-msr} in the Results section.

@@

I built a two-way fixed effect model in a difference-in-differences (DID) design, following Bai and colleagues' (2018; 2020) approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. I first specified an unrestricted and unadjusted model: \[\Delta score_{i,j} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i;1} + \lambda \cdot S_{j} + \varepsilon_{i,j}\] For student \(i\) in school \(j\), \(\Delta score_{i,j}\) is the change in test score between the two waves of surveys, \(migr_{i}\) is the treatment dummy variable, \(score_{i,j;1}\) is the test score at baseline, and \(S_{j}\) is the school fixed-effect. School fixed-effect can account for differences due to time-invariant factors at the school and county levels, as students nested in one school are also in the same county. An alternative is to use a class fixed-effect to control for time-invariant factors at the classroom level and above. However, about one-sixth of the students were reassigned into new classes at the end of baseline academic year. As the reassignment of classroom will complicate the model, I ruled out the classroom fixed-effect.

The DID model is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed effect model (Huntington-Klein, 2022). The standard errors are clustered at the classroom level.

In addition, I made an unrestricted and adjusted DID model: \[\Delta score_{i,j} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i;1} + \lambda \cdot S_{j} + \theta \cdot X_{i} + \varepsilon_{i,j}\] where \(X_{i}\) is the vector of covariates capturing the characteristics of children, their parents, and their households. The control variables were measured in the baseline survey. This version of model lifts the restriction that covariates from the baseline survey would be associated with a coefficient that equals one.

I adopted \texttt{R} (version 4.2.0) for all the analyses (R Core Team, 2022). This involved the following packages: \texttt{tidyverse} (version 1.3.1) for data cleaning (Wickham et al., 2019); \texttt{fixest} (version 0.10.4) for econometric modeling (Berge et al., 2022); \texttt{MatchIt} (version 4.4.0) and \texttt{cobalt} (version 4.3.2) for matching (Greifer, 2022b; Ho et al., 2022); \texttt{modelsummary} (version 1.0.0), \texttt{vtable} (version 1.3.3), and \texttt{kableExtra} (version 1.3.4) for presenting the results (Arel-Bundock, 2022; Huntington-Klein, 2021; Zhu, 2021); and \texttt{papaja} (version 0.1.0) for typesetting (Aust \& Barth, 2022).
